Score-based tests of measurement invariance: Use in practice

### 2014-06-30 frontiers/IMPSpres-knitr.Rnw
Revised imps presentation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@347 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\title{Score-based tests of measurement invariance: Use in practice}
\subtitle{2014 Annual Meeting of Psychometric Society}
\author[shortname]{Ting Wang\inst{1} \and Edgar C. Merkle\inst{1} \and 
  Achim Zeileis\inst{2}}
\institute[]{\inst{1} University of Missouri \and 
                      \inst{2} Universit\"{a}t Innsbruck}

@@

@@

%--------------------------------------------------------------------------
\begin{frame}
\frametitle{Measurement Invariance}
  \begin{itemize}
    \item Measurement invariance: Sets of tests/items consistently
      assigning scores across diverse groups of individuals.\\ \ \\
    \item Notable examination of measurement invariance:
      \begin{itemize}
         \item SAT for different ethnic groups \cite{atk01}
         \item Intelligence tests for different cohorts. \cite{wich04}
      \end{itemize}
  \end{itemize}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Example: Grade 1, 2, 3 vs. Grade 4, 5, 6}
 \begin{figure}
    \label{fig:pathinmodeltheory}
\resizebox{8.0cm}{!}{
\begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=20mm, minimum height=10mm},
 manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
                inner sep=0pt,minimum width=20mm, minimum height=10mm},
 error/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=10mm},
 post/.style={->,shorten >=1pt,>=stealth',semithick},
 postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
 postdashed/.style={->,shorten >=1pt,>=stealth',semithick,red,dashed},
 sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=20mm,semithick},
 errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=10mm,semithick}]

@@

\end{frame}
%--------------------------------------------------------------------------
%\begin{frame}
%\frametitle{Problem Scenario}
%\begin{itemize}
%  \item
% \emph{Data:} \\
% The data consist of responses to a gratitude scale from $n=1401$ youth 
% aged 10--19 years, dividing into six age groups: 10-11, 12-13, 14,
% 15, 16, 17-19. 
%  \item
%\emph{Question:} \\
%The researcher is interested in whether the scale
%is measurement invariant across these age groups.  
%\end{itemize}
%
%\end{frame}
%%--------------------------------------------------------------------------
%\begin{frame}
%  \frametitle{Similar Problem Scenarios and Generalization}
%  \begin{itemize}
%    \item Whether model parameter(s) hold(s) for people with different
%      income level, grade level, motivation level, etc.  
%    \item In general, the problem is whether measurement invariance(M.I.)
%      hypothesis holds with respect to an auxiliary ordinal variable. 
%    \end{itemize}
%\end{frame}
%%--------------------------------------------------------------------------
\begin{frame}
\frametitle{Possible Solution}
\begin{itemize}
\item Likelihood Ratio Test (LRT) (Require estimates under null
  hypothesis and alternative).
%  \begin{itemize}
%    \item Ignore the nature of ordinal variable, treat it as
%      categorical varaible. 
%    \item Sensitive to sample size. Rejecte the reduced model(s) too often.
%    \item Rather tedious and time consuming to fit and compare all possible
%      models. e.g, for testing one parameter, you need to fit $2^{m-1}$ 
%      models, $m$ is the level of ordinal variable.
%    \end{itemize}
\item Wald Test (Require estimates under alternative). 
\item Lagrange Multiplier Test (Score Test) (Require estimates under
  null hypotheis)
\item \textbf{Score-based Tests (more general) w.r.t. auxiliary
    variable (categorical, ordinal, continuous)}. 
\end{itemize}
\end{frame}
%  --------------------------------------------------------------------------
\begin{frame}
\frametitle{Outline} 
% Table of contents slide, comment this block out to remove it
\tableofcontents 
% Throughout your presentation, if you choose to use \section{} and 
%\subsection{} commands, these will automatically be printed on this slide 
%as an overview of your presentation
\end{frame}
%  --------------------------------------------------------------------------
\section{Theoretical Background}
\begin{frame}
  \frametitle{Score-based Tests}
  \begin{itemize}
    \item The score-based tests \cite{MerZei13, MerFanZei}
      rely on first derivatives of
      the model's log-likelihood function.
    \item We consider individual terms ({\em scores}) of the gradient.  
      These scores tell us how well a particular parameter describes a 
      particular individual.
  % Equation of gradient vs scores for individuals
  \end{itemize}

@@

\end{frame}
%--------------------------------------------------------------------------

\begin{frame}
  \frametitle{Empirical Cumulative Sum Score:  Systematic Order and 
    Decorrelate}
 We need a way to aggregate scores across people so that we
      can draw some general conclusions.
      \begin{itemize}
        \item Order individuals by an auxiliary variable.
        \item Decorrelate sores for each parameter.    
      \end{itemize}
\begin{equation} 
    \label{eq:cumscore}
  {\bm B}(t; \hat {\bm \theta}) ~=~ \hat {\bm I}^{-1/2} n^{-1/2}
    \sum_{i = 1}^{\lfloor n \cdot t \rfloor} {\bm s}(\hat {\bm \theta}; x_{(i)})
  \qquad (0 \le t \le 1)
\end{equation}

@@

\begin{frame}
  \frametitle{Convergence Distribution}
  \begin{itemize}
   \item  Under the hypothesis of measurement invariance, a central limit 
     theorem can be
   used to show that the fluctuation of the above empirical cumulative sum 
   score
   matrix ${\bm B}(\cdot; \hat {\bm \theta})$ converges in distribution to a 
   k-dimensional Brownian bridge ${\bm B}^{0}(\cdot)$.
   \begin{equation}
     \label{eq:converge}
    {\bm B}(\cdot; \hat {\bm \theta}) ~\overset{d}
        {\rightarrow}~ {\bm B}^{0}(\cdot),
   \end{equation}
   \item This result allows us to calculate $p$-values and
   critical values for test statistics under the hypothesis of
   measurement invariance.
  \end{itemize}
 \end{frame}
%--------------------------------------------------------------------------  
 \begin{frame}
   \frametitle{Statistics Based on Empirical Cumulative Sum Socre}
\begin{itemize}
  \item Statistics for auxiliary categorical variable:\\
maximum Lagrange multiplier categorical ($\mathit{LM}_{uo}$), 
asymptotic equvialent to LRT.\\
\item Statistics for auxiliary ordinal variable: \\
Weighted Double Maximum ordinal ($\mathit{WDM}_o$),\\
maximum Lagrange multiplier ordinal ($\max\mathit{LM}_o$). \\
\item Statistics for auxiliary continuous variable:\\
Double maximumum statistic ($\mathit{DM}$)\\
Cram{\'e}r-von Mises type statistic ($\mathit{CvM}$)\\
maximum Lagrange multiplier ($\max\mathit{LM}$)\\
\end{itemize}
\end{frame}

@@

%-------------------------------------------------------------------------   
\section{Simulation}
\begin{frame}
  \frametitle{Study Motivation}
  \begin{block}{Practice Consideration}
    \begin{itemize}
      \item Model misspecification is inevitable in practice.  
      \item M.I. is often studied with respect to \textbf{ordinal} variable.
      \item Availability/Convenience of software implementation.
    \end{itemize}
  \end{block}
\end{frame}
%-------------------------------------------------------------------------   
\begin{frame}
  \frametitle{Simulation Motivation}
  \begin{block}{Simulation:}
    What is the ordinal tests' power under model misspecification? 
    \begin{itemize}
      \item The estimated model excludes one parameter, which is \emph
        {not} the problematic parameter.
      \item The estimated model excludes one parameter, which \emph{is}
        the problematic parameter. 
    \end{itemize}
  \end{block}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Simulation Setup}
 \begin{itemize}
   \item Three sample sizes ($n=120,\textcolor{red}{480},960$), 
   \item Three numbers of ordinal levels ($m=4,8,12$), 
   \item Violation magnitude: $d$ times the parameters' asymptotic 
     standard errors (scaled by $\sqrt n$), with $d=0,0.25,0.5,...,4$ 
   \item Ordinal Violation: began at level $1+m/2$ of the auxiliary 
     variable $V$.
 \end{itemize}
   Statistics $\mathit{WDM}_o$, $\max \mathit{LM}_o$, $\mathit{LM}_{uo}$ were 
   examined.
\end{frame}
%--------------------------------------------------------------------------

\begin{frame}
  \begin{figure}
    \caption{Excluded out parameter is not the problematic parameter}
    \label{fig:pathinmodel1}
\resizebox{8.0cm}{!}{
\begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=20mm, minimum height=10mm},
 manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
                inner sep=0pt,minimum width=20mm, minimum height=10mm},
 error/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=10mm},
 post/.style={->,shorten >=1pt,>=stealth',semithick},
 postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
 postdashed/.style={->,shorten >=1pt,>=stealth',semithick,dashed},
 sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=20mm,semithick},
 errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=10mm,semithick}]

@@

\begin{frame}
\begin{figure}
\tiny{\caption{M.I.violation parameter $\lambda_{11}$.  Testing
    parameters are $\lambda_{11}$ and $\lambda_{11} \ldots \lambda_{62}$.}}
\label{fig:sim21res}
<<sim21res, fig.height=5, fig.width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra+loading" & 
          pars %in% c("Loading1", "All Loadings") & diff %% 0.5 == 0),
       type = "b", xlab = 
          expression(paste("Violation Magnitude (", lambda[11], ")")),
       ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
           if(which.given == 2){
             strip.default(which.given, factor.levels = 
                           parlabs[c(1, 5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\begin{figure}
\tiny{\caption{M.I.violation parameter $\lambda_{11}$.  Testing
    parameters are $\phi_{12}$, $\psi_{11} \ldots \psi_{66}$, and
    $\mu_{11} \ldots \mu_{66}$.}}
\label{fig:sim22res}
<<sim22res, fig.height=5, fig.width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra+loading" & 
          pars %in% c("All Errors", "Covariance", "All Intercepts") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = 
          expression(paste("Violation Magnitude (", lambda[11], ")")),
       ylab = "Power", ylim = c(0,1), key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
           if(which.given == 2){
             strip.default(which.given, factor.levels = 
                           parlabs[c(2, 6, 7)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

@@

\begin{frame}
\begin{figure}
    \caption{Excluded parameter is the problematic parameter}
    \label{fig:pathinmodel2}
\resizebox{8.0cm}{!}{
\begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=20mm, minimum height=10mm},
 manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
                inner sep=0pt,minimum width=20mm, minimum height=10mm},
 error/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=10mm},
 post/.style={->,shorten >=1pt,>=stealth',semithick},
 postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
 postdashed/.style={->,shorten >=1pt,>=stealth',semithick,red,dashed},
 sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=20mm,semithick},
 errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=10mm,semithick}]

@@

\begin{frame}
\begin{figure}
\tiny{\caption{The parameter violating measurement invariance is the 
    unmodeled loading. Testing parameters are $\lambda_{11}$ and $\psi_{11}$.}}
\label{fig:sim23res}
<<sim23res, fig.height=5, fig.width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra" & 
	 pars %in% c("Loading1", "Error1") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude (Unmodeled Loading)",
       ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	 strip.default(which.given, factor.levels = parlabs[c(1, 3)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@
\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\begin{figure}
\tiny{\caption{The parameter violating measurement invariance is 
    the unmodeled loading. Testing parameters are $\phi_{12}$ 
    and $\mu_{11}$.}}
\label{fig:sim24res}
<<sim24res, fig.height=5, fig.width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra" &
	 pars %in% c("Covariance", "Intercept1") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude (Unmodeled Loading)",
       ylab = "Power", ylim = c(0,1), key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	 strip.default(which.given, factor.levels = parlabs[c(2, 4)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@
\end{figure}
\end{frame}

@@

\begin{frame}
  \frametitle{Conclusion based on Simulation Studies}
 \begin{itemize}
   \item
      The ordinal test statistic proposed have better power 
      for detecting measurement 
      invariance violation parameter specifically that are monotonic 
      with the auxiliary ordinal variable. 
   \item This feature is robust as long as the violation 
      parameter is included in the testing model.
   \item The use of these tests is convenient without changing the 
      focal psychometric model.
   \end{itemize}
\end{frame}
%--------------------------------------------------------------------------
\section{Application}
\begin{frame}
  \frametitle{Application: Software Preparation}
The tests can be carried out in R. Specifically, 
\begin{itemize}
\item Model estimation can be carried out in \emph{lavaan} package. 
\item The test statistic and plot is got from \emph{strucchange} package.
\end{itemize}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[1.]
 \item Input data
\end{enumerate}
\begin{itemize}
  \item Gratitude scale (GQ-6) completed by six age
    groups(10-11, 12-13, 14, 15, 16, 17-19) \cite{frofan11}
  \item sample size is 1327.
  \item LRT is always significant when sample size is large. 
\end{itemize}
\end{frame}

@@

\end{frame}
%--------------------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[3.]
  \suite
  \item Test if M.I. is violated.
\end{enumerate}
<<sctest-ord2, eval=TRUE>>=
 library(strucchange)
wdmo <- sctest(restr, order.by = yg$agegroup, parm = 1:4, 
               vcov = "info", functional = "WDMo", 
               plot = FALSE)
wdmo
@ 
\end{frame}
%--------------------------------------------------------------------------

>>>>

\title{Score-based tests of measurement invariance: Use in practice}
\subtitle{2014 Annual Meeting of Psychometric Society}
\author[shortname]{Ting Wang\inst{1} \and Edgar C. Merkle\inst{1} \and 
  Achim Zeileis\inst{2}}
\institute[]{\inst{1} University of Missouri \and 
                      \inst{2} Universit\"{a}t Innsbruck}
\date{Supported by NSF grant SES-1061334}

@@

\renewcommand\bibliographytypesize{\tiny}

@@

%--------------------------------------------------------------------------
\begin{frame}
\frametitle{Measurement Invariance}
  \begin{itemize}
    \item Measurement invariance: Sets of tests/items consistently
      assigning scores across diverse groups of individuals.\\ \ \\
    \item Applications illustrating the importance of measurement invariance:
      \begin{itemize}
         \item SAT for different ethnic groups \cite{atk01}
         \item Intelligence tests' use over time. \cite{wich04}
      \end{itemize}
  \end{itemize}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Example: Grade 5 vs. Grade 6}
 \begin{figure}
    \label{fig:pathinmodeltheory}
\resizebox{8.0cm}{!}{
\begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=20mm, minimum height=10mm},
 manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
                inner sep=0pt,minimum width=20mm, minimum height=10mm},
 error/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=10mm},
 post/.style={->,shorten >=1pt,>=stealth',semithick},
 postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
 postdashed/.style={->,shorten >=1pt,>=stealth',semithick,red,dashed},
 sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=20mm,semithick},
 errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=10mm,semithick}]

@@

\end{frame}
%--------------------------------------------------------------------------
%\begin{frame}
%\frametitle{Problem Scenario}
%\begin{itemize}
%  \item
% \emph{Data:} \\
% The data consist of responses to a gratitude scale from $n=1401$ youth 
% aged 10--19 years, dividing into six age groups: 10-11, 12-13, 14,
% 15, 16, 17-19. 
%  \item
%\emph{Question:} \\
%The researcher is interested in whether the scale
%is measurement invariant across these age groups.  
%\end{itemize}
%
%\end{frame}
%%--------------------------------------------------------------------------
%\begin{frame}
%  \frametitle{Similar Problem Scenarios and Generalization}
%  \begin{itemize}
%    \item Whether model parameter(s) hold(s) for people with different
%      income level, grade level, motivation level, etc.  
%    \item In general, the problem is whether measurement invariance(M.I.)
%      hypothesis holds with respect to an auxiliary ordinal variable. 
%    \end{itemize}
%\end{frame}
%%--------------------------------------------------------------------------
\begin{frame}
\frametitle{Possible Solution}
\begin{itemize}
\item Likelihood Ratio Test (LRT): Require estimates under null
  hypothesis and alternative.
%  \begin{itemize}
%    \item Ignore the nature of ordinal variable, treat it as
%      categorical varaible. 
%    \item Sensitive to sample size. Rejecte the reduced model(s) too often.
%    \item Rather tedious and time consuming to fit and compare all possible
%      models. e.g, for testing one parameter, you need to fit $2^{m-1}$ 
%      models, $m$ is the level of ordinal variable.
%    \end{itemize}
\item Wald Test: Require estimates under alternative. 
\item Lagrange Multiplier Test (Score Test): Require estimates under
  null hypotheis.
\item \textbf{Score Test generalizations w.r.t. general auxiliary
    variable (categorical, ordinal, or continuous)}. 
\end{itemize}
\end{frame}
%  --------------------------------------------------------------------------
\begin{frame}
\frametitle{Outline} 
% Table of contents slide, comment this block out to remove it
\tableofcontents 
% Throughout your presentation, if you choose to use \section{} and 
%\subsection{} commands, these will automatically be printed on this slide 
%as an overview of your presentation
\end{frame}
%  --------------------------------------------------------------------------
\section{Theoretical Background}
\begin{frame}
  \frametitle{Score-based Tests}
  \begin{itemize}
    \item The score-based tests \cite{MerZei13, MerFanZei}
      rely on first derivatives of
      the model's log-likelihood function.
    \item We consider individual terms ({\em scores}) of the gradient.  
      These scores tell us how well a particular parameter describes a 
      particular individual.
  % Equation of gradient vs scores for individuals
  \end{itemize}

@@

\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Score Aggregation}
  \begin{itemize}
    \item Each individual has a score for each model parameter,
      resulting in $nk$ total scores (where $k$ is the dimension of
      $\bm{\theta}$).
    \item To draw general conclusions about measurement invariance, we
      need to aggregate across scores in specific manners.
    \item When $V$ is ordinal or continuous, we can make use of the
      ordering in $V$ to obtain novel test statistics.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Cumulative Sum of Scores}
      \begin{itemize}
        \item Order individuals by an auxiliary variable.
        \item Decorrelate scores, using an estimate of the information
          matrix.    
      \end{itemize}
\begin{equation} 
    \label{eq:cumscore}
  {\bm B}(t; \hat {\bm \theta}) ~=~ \hat {\bm I}^{-1/2} n^{-1/2}
    \sum_{i = 1}^{\lfloor n \cdot t \rfloor} {\bm s}(\hat {\bm \theta}; x_{(i)})
  \qquad (0 \le t \le 1)
\end{equation}

@@

\begin{frame}
  \frametitle{Distributional Theory}
  \begin{itemize}
   \item  Under the hypothesis of measurement invariance, a central limit 
     theorem can be
   used to show that the fluctuation of the above empirical cumulative sum 
   score
   matrix ${\bm B}(\cdot; \hat {\bm \theta})$ converges in distribution to a 
   k-dimensional Brownian bridge ${\bm B}^{0}(\cdot)$.
   \begin{equation}
     \label{eq:converge}
    {\bm B}(\cdot; \hat {\bm \theta}) ~\overset{d}
        {\rightarrow}~ {\bm B}^{0}(\cdot),
   \end{equation}
   \item This result allows us to calculate $p$-values and
   critical values for test statistics under the hypothesis of
   measurement invariance.
  \end{itemize}
 \end{frame}
%--------------------------------------------------------------------------  
 \begin{frame}
   \frametitle{Test Statistics}
\begin{itemize}
  \item Categorical $V$: Maximum Lagrange multiplier test (asymptotically equvialent to LRT).
\item Ordinal $V$: Weighted double maximum test ($\mathit{WDM}_o$);\\
Maximum Lagrange multiplier test ($\max\mathit{LM}_o$).
\item Continuous $V$: Double maximumum test ($\mathit{DM}$)\\
Cram{\'e}r-von Mises type statistic ($\mathit{CvM}$)\\
maximum Lagrange multiplier ($\max\mathit{LM}$)\\
\end{itemize}
\end{frame}

@@

%-------------------------------------------------------------------------   
\section{Simulation}
\begin{frame}
  \frametitle{Study Motivation}
  \begin{block}{Practical Considerations}
    \begin{itemize}
      \item Model misspecification is inevitable in practice.  
      \item Measurement invariance is often studied with respect to an
        \textbf{ordinal} variable.
      \item Availability/Convenience of software implementation.
    \end{itemize}
  \end{block}
\end{frame}
%-------------------------------------------------------------------------   
\begin{frame}
  \frametitle{Simulation Motivation}
  \begin{block}{Simulation:}
    What are the ordinal tests' power under model misspecification? 
    \begin{itemize}
      \item The estimated model excludes one parameter, which is \emph
        {not} the problematic parameter.
      \item The estimated model excludes one parameter, which \emph{is}
        the problematic parameter. 
    \end{itemize}
  \end{block}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Simulation Setup}
 \begin{itemize}
   \item Three sample sizes ($n=120,\textcolor{red}{480},960$), 
   \item Three numbers of ordinal levels ($m=4,8,12$), 
   \item Violation magnitude: $d$ times the parameters' asymptotic 
     standard errors (scaled by $\sqrt n$), with $d=0,0.25,0.5,...,4$ 
   \item Ordinal Violation: began at level $1+m/2$ of the auxiliary 
     variable $V$.
 \end{itemize}
   Statistics $\mathit{WDM}_o$, $\max \mathit{LM}_o$, $\mathit{LM}_{uo}$ were 
   examined.
\end{frame}
%--------------------------------------------------------------------------

\begin{frame}
  \begin{figure}
    \caption{Excluded parameter does not violate invariance}
    \label{fig:pathinmodel1}
\resizebox{8.0cm}{!}{
\begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=20mm, minimum height=10mm},
 manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
                inner sep=0pt,minimum width=20mm, minimum height=10mm},
 error/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=10mm},
 post/.style={->,shorten >=1pt,>=stealth',semithick},
 postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
 postdashed/.style={->,shorten >=1pt,>=stealth',semithick,dashed},
 sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=20mm,semithick},
 errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=10mm,semithick}]

@@

\begin{frame}
\begin{figure}
\tiny{\caption{$\lambda_{11}$ violates measurement invariance.  Tested
    parameters are $\lambda_{11}$ and \{$\lambda_{11}, \ldots, \lambda_{62}$.\}}}
\label{fig:sim21res}
<<sim21res, fig.height=5, fig.width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra+loading" & 
          pars %in% c("Loading1", "All Loadings") & diff %% 0.5 == 0),
       type = "b", xlab = 
          expression(paste("Violation Magnitude (", lambda[11], ")")),
       ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
           if(which.given == 2){
             strip.default(which.given, factor.levels = 
                           parlabs[c(1, 5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\begin{figure}
\tiny{\caption{$\lambda_{11}$ violates measurement invariance.  Tested
    parameters are $\phi_{12}$, \{$\psi_{11}, \ldots, \psi_{66}$\}, and
    \{$\mu_{11}, \ldots, \mu_{66}$\}.}}
\label{fig:sim22res}
<<sim22res, fig.height=5, fig.width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra+loading" & 
          pars %in% c("All Errors", "Covariance", "All Intercepts") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = 
          expression(paste("Violation Magnitude (", lambda[11], ")")),
       ylab = "Power", ylim = c(0,1), key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
           if(which.given == 2){
             strip.default(which.given, factor.levels = 
                           parlabs[c(2, 6, 7)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

@@

\begin{frame}
\begin{figure}
    \caption{Excluded parameter does violate invariance}
    \label{fig:pathinmodel2}
\resizebox{8.0cm}{!}{
\begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=20mm, minimum height=10mm},
 manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
                inner sep=0pt,minimum width=20mm, minimum height=10mm},
 error/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=10mm},
 post/.style={->,shorten >=1pt,>=stealth',semithick},
 postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
 postdashed/.style={->,shorten >=1pt,>=stealth',semithick,red,dashed},
 sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=20mm,semithick},
 errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=10mm,semithick}]

@@

\begin{frame}
\begin{figure}
\tiny{\caption{The parameter violating measurement invariance is the 
    unmodeled loading. Tested parameters are $\lambda_{11}$ and $\psi_{11}$.}}
\label{fig:sim23res}
<<sim23res, fig.height=5, fig.width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra" & 
	 pars %in% c("Loading1", "Error1") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude (Unmodeled Loading)",
       ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	 strip.default(which.given, factor.levels = parlabs[c(1, 3)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@
\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\begin{figure}
\tiny{\caption{The parameter violating measurement invariance is 
    the unmodeled loading. Tested parameters are $\phi_{12}$ 
    and $\mu_{11}$.}}
\label{fig:sim24res}
<<sim24res, fig.height=5, fig.width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra" &
	 pars %in% c("Covariance", "Intercept1") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude (Unmodeled Loading)",
       ylab = "Power", ylim = c(0,1), key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	 strip.default(which.given, factor.levels = parlabs[c(2, 4)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@
\end{figure}
\end{frame}

@@

\begin{frame}
  \frametitle{Conclusion based on Simulation Studies}
 \begin{itemize}
   \item
      The ordinal test statistics have better power 
      for detecting measurement 
      invariance violations that are monotonic 
      with $V$.
   \item This feature is robust as long as the violating 
      parameter is included in the testing model.
   \item The use of these tests is convenient, requiring estimation of
     only a reduced model.
   \end{itemize}
\end{frame}
%--------------------------------------------------------------------------
\section{Application}
\begin{frame}
  \frametitle{Software Illustration}
The tests can be carried out in R. Specifically, 
\begin{itemize}
\item Model estimation can be carried out in \emph{lavaan} package. 
\item The test statistic and plot is got from \emph{strucchange} package.
\end{itemize}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[1.]
 \item Input data
\end{enumerate}
\begin{itemize}
  \item Gratitude scale (GQ-6) completed by six age
    groups(10-11, 12-13, 14, 15, 16, 17-19) \cite{frofan11}
  \item $n=1327$.
  \item LRT is always significant when sample size is large. 
\end{itemize}
\end{frame}

@@

\end{frame}
%--------------------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[3.]
  \suite
  \item Test of measurement invariance
\end{enumerate}
<<sctest-ord2, eval=TRUE>>=
 library(strucchange)
wdmo <- sctest(restr, order.by = yg$agegroup, parm = 1:4, 
               vcov = "info", functional = "WDMo", 
               plot = FALSE)
wdmo
@ 
\end{frame}
%--------------------------------------------------------------------------

\begin{frame}
\begin{figure}
\tiny{\caption{Fluctuation processes for the $\mathit{WDM}_o$ statistic}}
\label{fig:ordres}
<<sctest-ord-graphics, fig.height=3.3, fig.width=5, echo=FALSE, results='hide'>>=
sctest(restr, order.by = yg$agegroup, parm = 1:4, vcov = "info",
  functional = "WDMo", plot = TRUE, 
               ylim = c(0, 3.2), xlab = "Age group", 
               ylab = "Weighted test statistics")
@
\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
 \begin{itemize}
 \item Thanks! 
 \item Questions?
 \end{itemize}
\end{frame}
%--------------------------------------------------------------------------
\appendix
\begin{frame}{Reference}
\bibliographystyle{apacite}
\bibliography{refs}
\end{frame}
%--------------------------------------------------------------------------
